Semantic Linearity in a Lazy Optimising Compiler

### 2024-02-13 icfp.tex
RM edits 3.4

% Indeed, usage environments allow us to encode mutual exclusivity between
% alternative ways of consuming linear resources (between $\D$-vars and direct
% resource usage). 
Let-bound variables are the canonical example of a
$\Delta$-bound variable, that is, let-bound variables bind expressions in which
the resources required to type them are consumed lazily rather than eagerly.
%
Annotating let-bound variables with a usage environment $\D$
\emph{delays} the consumption of resources to when the variables themselves are
used:
\[
\TypeLet
\]
The rule for (non-recursive) let bindings splits the linear environment in $\D$
and $\D'$. $\D$ is used to type the body $e$ of the let binding $x$. Perhaps
surprisingly, the resources $\D$ used to type $e$ are still available in the
environment to type the let body $e'$, alongside the unrestricted $x$ binding
annotated with the usage environment $\D$. Ultimately, the resources being
available in $e'$ reflects the fact that typing a lazily bound expression
doesn't consume resources, and the binding $x$ being $\D$-bound reflects that
its usage entails consuming the resources $\D$ the expression it binds depends on.
% \todo[inline]{Let bindings are hard, if they are used then we use resources. If
% they don't get used then we use no resources! In practice, resources that show
% up in the body of the let must be used, be it by using the let binder, or by
% using them directly. This makes the let binder and the resources in its body
% mutual exclusive.}

>>>>

% Indeed, usage environments allow us to encode mutual exclusivity between
% alternative ways of consuming linear resources (between $\D$-vars and direct
% resource usage). 
Let-bound variables are the canonical example of a
$\Delta$-bound variable.
% , that is, let-bound variables bind expressions in which
% the resources required to type them are consumed lazily rather than eagerly.
%
Annotating let-bound variables with a usage environment $\D$
\emph{delays} the consumption of resources to when the variables themselves are
used:
\[
\TypeLet
\]
The rule for (non-recursive) let bindings splits the linear environment in $\D$
and $\D'$. $\D$ is used to type the body $e$ of the let binding $x$. Perhaps
surprisingly, the resources $\D$ used to type $e$ are still available in the
environment to type the let body $e'$, alongside the unrestricted $x$ binding
annotated with the usage environment $\D$. In essence, the resources of $e$ being
available in $e'$ reflects how typing the body of a lazy binder
doesn't consume resources, and the binding $x$ being an unrestricted
$\D$-variable ensures that using it will consume the resources $\D$ the
expression it binds depends on.
% \todo[inline]{Let bindings are hard, if they are used then we use resources. If
% they don't get used then we use no resources! In practice, resources that show
% up in the body of the let must be used, be it by using the let binder, or by
% using them directly. This makes the let binder and the resources in its body
% mutual exclusive.}
